Erosion of somatic tissue identity with loss of the X-linked intellectual disability factor KDM5C

### 2024-07-26 submission/manuscript.Rmd
remove dazl protein from fig 2, put in functional section

We were particularly interested in the aberrant transcription of _Dazl_, since it is essential for germ cell development and promotes the translation of germline mRNAs[@liDAZLMasterTranslational2019]. A significant portion of germline transcripts misexpressed in _Kdm5c_-KO EpiLCs are known binding targets of DAZL, including _Stra8_[@mikedisDAZLMediatesBroad2020] (p = 1.698e-07, Fisher’s Exact Test). This suggests expression of DAZL protein could promote the translation of other aberrant germline transcripts, influencing their ability to impact _Kdm5c_-KO cellular function. We thus tested DAZL protein expression in _Kdm5c_-KO EpiLCs through immunocytochemistry (Figure 3I). We observed about 25% of _Kdm5c_-KO EpiLCs expressed DAZL protein and it was localized to the cytoplasm (p = 0.0015, Welch's t-test), consistent with the pattern of DAZL expression in spermatogonia[@mikedisDAZLMediatesBroad2020]. Altogether these results suggest tissue-specific genes are misexpressed during _Kdm5c_-KO embryogenesis, including key drivers of germline identity that can be translated into protein.

@@

\clearpage
# Notes
## Discussion notes

>>>>

@@

\clearpage
# Notes
## Dazl
We were particularly interested in the aberrant transcription of _Dazl_, since it is essential for germ cell development and promotes the translation of germline mRNAs[@liDAZLMasterTranslational2019]. A significant portion of germline transcripts misexpressed in _Kdm5c_-KO EpiLCs are known binding targets of DAZL, including _Stra8_[@mikedisDAZLMediatesBroad2020] (p = 1.698e-07, Fisher’s Exact Test). This suggests expression of DAZL protein could promote the translation of other aberrant germline transcripts, influencing their ability to impact _Kdm5c_-KO cellular function. We thus tested DAZL protein expression in _Kdm5c_-KO EpiLCs through immunocytochemistry (Figure 3I). We observed about 25% of _Kdm5c_-KO EpiLCs expressed DAZL protein and it was localized to the cytoplasm (p = 0.0015, Welch's t-test), consistent with the pattern of DAZL expression in spermatogonia[@mikedisDAZLMediatesBroad2020]. Altogether these results suggest tissue-specific genes are misexpressed during _Kdm5c_-KO embryogenesis, including key drivers of germline identity that can be translated into protein.

## Discussion notes
